---
title: "Statistical Entropy Analysis of Network Data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Statistical Entropy Analysis of Network Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(out.width = "100%",
  cache = FALSE
)
```


This vignette serves as an introduction to the package `netropy`. Parts of the theoretical background is provided but for more details, consult the following literature which the package is based on:

> Frank, O., & Shafie, T. (2016). Multivariate entropy analysis of network data.
*Bulletin of Sociological Methodology/Bulletin de Méthodologie Sociologique*, 129(1), 45-63.

>  Nowicki, K., Shafie, T., & Frank, O. (Forthcoming 2022). *Statistical Entropy Analysis of Network Data*.

Make sure the library is loaded
```{r eval=TRUE}
library('netropy')
```


## Introduction


### Loading Internal Data

### Variable Domains and Data Editing

## Univariate, Bivariate and Trivariate Entropies

## Joint Entropies and Association Graphs



## Prediction Powers Based on Expected Conditional Entropies

